---
title: "ClinicalDataTable"
author: "Natalie Geyer"
date: "4/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages
```{r}
library(rtables)
library(tibble)
library(knitr)
library(report)
library(dplyr)
```

## Clinical Data Summary Table

### Set-Up
```{r, echo=FALSE}
# Import directory for entire dataframe (sorted by Probe)
clinPatAnnotFn <- "./output/clin_data_annot.csv"
```

Load entire datafram with clinical and pathological data. 

```{r, echo=TRUE}
clinPatAnnotData <- read.csv(clinPatAnnotFn)

clinPatAnnotData_l <- clinPatAnnotData %>% 
  mutate(sex = recode(sex, m = "male")) %>% 
  mutate(sex = recode(sex, f = "female"))
```

### Build the table 
We want the following variables in the table: age, sex, primary_side, sync_meta, num_tumors, marginal1, max_diameter, sum_diameter, tum_percent_annot
for all patients and sorted for neoadjuvant, chemonaive patients
```{r, echo=FALSE}
s_summary <- function(x) {
  if (is.numeric(x)) {
    in_rows(
      "n" = rcell(sum(!is.na(x)), format = "xx"),
      "Mean (sd)" = rcell(c(mean(x, na.rm = TRUE), sd(x, na.rm = TRUE)), format = "xx.xx (xx.xx)"),
      "IQR" = rcell(IQR(x, na.rm = TRUE), format = "xx.xx"),
      "min - max" = rcell(range(x, na.rm = TRUE), format = "xx.xx - xx.xx")
    )
  } else if (is.factor(x)) {
    
    vs <- as.list(table(x))
    do.call(in_rows, lapply(vs, rcell, format = "xx"))
    
  } else (
    stop("type not supported")
  )
}

lyt <- basic_table() %>% 
  split_cols_by(var = "neoadjuvant") %>%
  add_colcounts() %>%
  analyze(c("age", "sex", "primary_side", "sync_meta", "num_tumors", "marginal1", "max_diameter", "sum_diameter", "tum_percent_annot"), afun = s_summary) 

tbl <- build_table(lyt, clinPatAnnotData_l)
tbl
```

## Report the source of applied packages
```{r}
cite_packages()
```