---
title: "ClinicalDataTable"
author: "Natalie Geyer"
date: "4/20/2022"
output:
  pdf_document: 
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages
```{r, warning=FALSE, message=FALSE}
library(rtables)
library(finalfit)
library(tibble)
library(knitr)
library(report)
library(dplyr)
library(ggplot2)
```

## Clinical Data Summary Table

### Set-Up
```{r, echo=TRUE}
# Import directory for entire dataframe (sorted by Probe)
clinPatAnnotFn <- "./output/clin_data_annot.csv"
```

Load entire datafram with clinical and pathological data. 

```{r, echo=TRUE}
clinPatAnnotData <- read.csv(clinPatAnnotFn)

clinPatAnnotData_l <- clinPatAnnotData %>% 
  mutate(sex = recode(sex, m = "male")) %>% 
  mutate(sex = recode(sex, f = "female"))
```

### Build the table, version 1 with rtables package 
We want the following variables in the table: age, sex, primary_side, sync_meta, num_tumors, marginal1, max_diameter, sum_diameter, tum_percent_annot
for all patients and sorted for neoadjuvant, chemonaive patients
```{r, echo=FALSE}
s_summary <- function(x) {
  if (is.numeric(x)) {
    in_rows(
      "n" = rcell(sum(!is.na(x)), format = "xx"),
      "Mean (sd)" = rcell(c(mean(x, na.rm = TRUE), sd(x, na.rm = TRUE)), format = "xx.xx (xx.xx)"),
      "IQR" = rcell(IQR(x, na.rm = TRUE), format = "xx.xx"),
      "min - max" = rcell(range(x, na.rm = TRUE), format = "xx.xx - xx.xx")
    )
  } else if (is.factor(x)) {
    
    vs <- as.list(table(x))
    do.call(in_rows, lapply(vs, rcell, format = "xx"))
    
  } else (
    stop("type not supported")
  )
}

lyt <- basic_table() %>% 
  split_cols_by(var = "neoadjuvant") %>%
  add_colcounts() %>%
  analyze(c("age", "sex", "primary_side", "sync_meta", "num_tumors", "marginal1", "max_diameter", "sum_diameter", "tum_percent_annot"), afun = s_summary) 

tbl <- build_table(lyt, clinPatAnnotData_l)
tbl
```

### Build the table, version 2 with finalfit package 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# primary_side, sync_meta, num_tumors, marginal1, max_diameter, sum_diameter, tum_percent_annot
# BUG max_diameter hat 19mal "-Inf" als Wert

#define dependent variable
dependent <- "neoadjuvant"

#define explanatory variables 
clinPatAnnotData_l2 <- clinPatAnnotData_l %>% 
  select(age, 
         sex, 
         primary_side, 
         sync_meta, 
         num_tumors, 
         marginal1, 
         #max_diameter, 
         sum_diameter, 
         tum_percent_annot 
         ) %>% 
  names() -> explanatory

clinPatAnnotData_l %>% 
  ff_glimpse(dependent, explanatory)

clinPatAnnotData_l %>% 
  summary_factorlist(dependent, explanatory, 
    p=TRUE, na_include=TRUE, 
    add_dependent_label=TRUE) -> table1

kable(table1)

# analyze correlation with status_fu_os (overall survival, patient dead yes/no bzw 1/0)
# explanatory variables defined above

#define dependent variable
clinPatAnnotData_l$status_fu_os <- as.factor(clinPatAnnotData_l$status_fu_os)
dependentOS = "status_fu_os"

#build table
clinPatAnnotData_l %>% 
  finalfit(dependentOS, explanatory) -> table2
kable(table2)

clinPatAnnotData_l %>% 
  or_plot(dependentOS, explanatory, 
  column_space = c(-0.5, 0, 0.5),
  table_text_size = 3,
  title_text_size = 10,
  plot_opts=list(xlab("Odds ratio (95% CI, log scale)"), theme(axis.title = element_text(size=10)))
  )
  
save(table1, table2, dependent, dependentOS, explanatory, file = "out.rda")


```


## Report the source of applied packages
```{r}
cite_packages()
```