---
title: "Visual scoring (EvelinaDanyil) vs annotation-based scoring_review"
author: "Natalie Geyer"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generated on: `r Sys.time()`

Import packages
```{r}
#Test commit changes 2022-04-19
library(readxl)
library(tidyverse)
library(irr)
library(ggExtra)
library(ggplot2)
library(knitr)
```


```{r}
PeterVEstFn <-  "./input/Data visual scorings/Teaching_set_HGP_scores_PV_bis.xlsx" # Only 1-104
DanyilEstFn <- "./input/Data visual scorings/HGP_scores_Danyil.xlsx" # Only 59-104
EvelinaEstFn <- "./input/Data visual scorings/Template_and_example_HGP_scoresEVELINA.xlsx" # Only 1-58
KSSlideFn <- "./output/gp_annotations_by_slide.csv" 
```


Import and initialize KS data
```{r}
# Obtain KS data in tabulated form
KSSlideDataTab <- read.csv(KSSlideFn, row.names=NULL)
KSSlideDataTab <- KSSlideDataTab %>% mutate(slide_name = paste(ids, tumors, blocks, sep="-")) %>% relocate(slide_name)
KSSlideDataTab$slide_name <- as.factor(KSSlideDataTab$slide_name)
head(KSSlideDataTab)

# Recode GP names and R2 to R
KSSlideDataLong <- KSSlideDataTab %>% select(-tumors, -blocks)
KSSlideDataLong <- KSSlideDataLong %>% mutate(annotation_types = recode(annotation_types, R2 = "R"))
KSSlideDataLong <- KSSlideDataLong %>% mutate(annotation_types = recode(annotation_types, D = "desmo"))
KSSlideDataLong <- KSSlideDataLong %>% mutate(annotation_types = recode(annotation_types, R = "replacement"))
KSSlideDataLong <- KSSlideDataLong %>% mutate(annotation_types = recode(annotation_types, P = "pushing"))

# Initialize % of each GP for every slide, so we can compare all and avoid lack of scoring for some slides
for(an_slid in unique(KSSlideDataLong$slide_name)) {
  an_id <- as.integer(unlist(strsplit(an_slid, "-"))[1])
  KSSlideDataLong <- KSSlideDataLong %>% add_row(slide_name = an_slid, ids = an_id, annotation_types = factor("desmo"), percent_gp = 0, length_um = 0)
  KSSlideDataLong <- KSSlideDataLong %>% add_row(slide_name = an_slid, ids = an_id, annotation_types = factor("replacement"), percent_gp = 0, length_um = 0)
  KSSlideDataLong <- KSSlideDataLong %>% add_row(slide_name = an_slid, ids = an_id, annotation_types = factor("pushing"), percent_gp = 0, length_um = 0)
}

# Summarize again (sum up) GP values
KSSlideDataLong <- KSSlideDataLong %>% group_by(ids, slide_name, annotation_types) %>% summarize(percent_gp = sum(percent_gp), length_um = sum(length_um))
dim(KSSlideDataLong)
length(unique(KSSlideDataLong$slide_name))
head(KSSlideDataLong)

# KS slide data
KSSlideGPDataLong <- KSSlideDataLong %>% mutate(slide_name = paste0(slide_name, '-', annotation_types)) %>% ungroup() %>% select(-ids, -annotation_types, -length_um) %>% rename(percent_gp_KS = "percent_gp")
dim(KSSlideGPDataLong)
length(unique(KSSlideGPDataLong$slide_name))/3
head(KSSlideGPDataLong)

# Transfor KS data to longer format to match ...EstSlideDataWide
KSSlideDataWide <- KSSlideDataLong %>% pivot_wider(id_cols = slide_name, names_from = annotation_types, values_from = percent_gp, values_fill = list(percent_gp = 0))
head(KSSlideDataWide)
```


```{r, include=FALSE}
# Import and initialize Peter data 

# Obtain Peter data
PeterVEstSlideDataWide <- read_excel(PeterVEstFn, col_types = c("text", rep("numeric", 3), rep("skip",4)))
PeterVEstSlideDataWide <- rename(PeterVEstSlideDataWide, slide_name = "Code", desmo = "D", replacement = "R", pushing = "P")
head(PeterVEstSlideDataWide)

# Pivot to long format
PeterVEstSlideGPDataLong <- PeterVEstSlideDataWide %>% pivot_longer(-slide_name, names_to = "annotation_types", values_to = "percent_gp")
PeterVEstSlideGPDataLong <- PeterVEstSlideGPDataLong %>% mutate(slide_name = paste0(slide_name, '-', annotation_types)) %>% select(-annotation_types) %>% rename(percent_gp_Peter = "percent_gp")
dim(PeterVEstSlideGPDataLong)
length(unique(PeterVEstSlideGPDataLong$slide_name))/3
head(PeterVEstSlideGPDataLong)
```

Import and initialize Evelina/Danyil data
```{r}
# Obtain Danyil and Evelina data and remove empty rows

EvelinaEstSlideDataWide <- read_excel(EvelinaEstFn, col_types = c("text", rep("numeric", 3), rep("skip",4)))
EvelinaEstSlideDataWide <- rename(EvelinaEstSlideDataWide, slide_name = "Code", desmo = "D", replacement = "R", pushing = "P")
EvelinaEstSlideDataWide <- EvelinaEstSlideDataWide[!is.na(EvelinaEstSlideDataWide$desmo),]
EvelinaEstSlideDataWide <- EvelinaEstSlideDataWide[!with(EvelinaEstSlideDataWide,is.na(EvelinaEstSlideDataWide$desmo) & is.na(EvelinaEstSlideDataWide$replacement) & is.na(EvelinaEstSlideDataWide$pushing)),] # remove empty rows
head(EvelinaEstSlideDataWide)

DanyilEstSlideDataWide <- read_excel(DanyilEstFn, col_types = c("text", rep("numeric", 3), rep("skip",4)))
DanyilEstSlideDataWide <- rename(DanyilEstSlideDataWide, slide_name = "Code", desmo = "D", replacement = "R", pushing = "P")
DanyilEstSlideDataWide <- DanyilEstSlideDataWide[!with(DanyilEstSlideDataWide,is.na(DanyilEstSlideDataWide$desmo) & is.na(DanyilEstSlideDataWide$replacement) & is.na(DanyilEstSlideDataWide$pushing)),] # remove empty rows
head(DanyilEstSlideDataWide)

# Combine 

EDEstSlideDataWide <- rbind(EvelinaEstSlideDataWide, DanyilEstSlideDataWide)
head(EDEstSlideDataWide)

# Pivot to long format

EDEstSlideDataLong <- EDEstSlideDataWide %>% pivot_longer(-slide_name, names_to = "annotation_types", values_to = "percent_gp")
EDEstSlideDataLong <- EDEstSlideDataLong %>% mutate(slide_name = paste0(slide_name, '-', annotation_types)) %>% select(-annotation_types) %>% rename(percent_gp_EvelinaDanyil = "percent_gp")
dim(EDEstSlideDataLong)
length(unique(EDEstSlideDataLong$slide_name))/3
head(EDEstSlideDataLong)
```

```{r, include=FALSE}
# Merged KS-Peter scores for every slide and GP

# Merge scores
KSPeterSlideGPDataLong <- merge(KSSlideGPDataLong, PeterVEstSlideGPDataLong)
dim(KSPeterSlideGPDataLong)
length(unique(KSPeterSlideGPDataLong$slide_name))/3
head(KSPeterSlideGPDataLong)

# QA: Check for discrepancies in scored slides
QAKSPeterSlideGPDataLong <- merge(KSSlideGPDataLong, PeterVEstSlideGPDataLong, all = T)
dim(QAKSPeterSlideGPDataLong)
length(unique(QAKSPeterSlideGPDataLong$slide_name))/3
#  probes with nas
unique(QAKSPeterSlideGPDataLong %>% filter(is.na(percent_gp_Peter) & is.na(percent_gp_KS)) %>% pull(slide_name))
unique(QAKSPeterSlideGPDataLong %>% filter(is.na(percent_gp_Peter) & !is.na(percent_gp_KS)) %>% pull(slide_name))
unique(QAKSPeterSlideGPDataLong %>% filter(!is.na(percent_gp_Peter) & is.na(percent_gp_KS)) %>% pull(slide_name))

# Visualize, scatter+histogram
ScatterKSPeterV <- ggplot(KSPeterSlideGPDataLong, aes(x = percent_gp_Peter, y = percent_gp_KS)) + geom_point() + theme_classic()
ggExtra::ggMarginal(ScatterKSPeterV, type = "histogram", xparams = list(bins = 20, fill = "blue"), yparams = list(bins = 100, fill = "orange"))

```

Merged KS-Evelina/Danyil scores for every slide and GP
```{r}
# Merge scores
KSEDSlideGPDataLong <- merge(KSSlideGPDataLong, EDEstSlideDataLong)
dim(KSEDSlideGPDataLong)
length(unique(KSEDSlideGPDataLong$slide_name))/3
head(KSEDSlideGPDataLong)

# QA: Check for discrepancies in scored slides
QAKSEDSlideGPDataLong <- merge(KSSlideGPDataLong, EDEstSlideDataLong, all = T)
dim(QAKSEDSlideGPDataLong)
length(unique(QAKSEDSlideGPDataLong$slide_name))/3
#  probes with nas
unique(QAKSEDSlideGPDataLong %>% filter(is.na(percent_gp_EvelinaDanyil) & is.na(percent_gp_KS)) %>% pull(slide_name))
unique(QAKSEDSlideGPDataLong %>% filter(is.na(percent_gp_EvelinaDanyil) & !is.na(percent_gp_KS)) %>% pull(slide_name))
unique(QAKSEDSlideGPDataLong %>% filter(!is.na(percent_gp_EvelinaDanyil) & is.na(percent_gp_KS)) %>% pull(slide_name))
```

Visualization visual Annotation vs whole slide image Annotation Evelina&Danyil and save as pdf output
```{r}
# Visualize and save as pdf, scatter+histogram
# Step 1: Call the pdf command to start the plot
pdf(file = "./plots/ScatterHistoVisualEDvsWSI.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 6) # The height of the plot in inches
# Step 2: Create the plot with R code
ScatterKSED <- ggplot(KSEDSlideGPDataLong, aes(x = percent_gp_EvelinaDanyil, y = percent_gp_KS)) + geom_point() + theme_classic()
ggExtra::ggMarginal(ScatterKSED, type = "histogram", size = 3, xparams = list(bins = 20, fill = "blue"), yparams = list(bins = 100, fill = "orange"))
# Step 3: Run dev.off() to create the file!
dev.off()

ScatterKSED <- ggplot(KSEDSlideGPDataLong, aes(x = percent_gp_EvelinaDanyil, y = percent_gp_KS)) + geom_point() + theme_classic()
ggExtra::ggMarginal(ScatterKSED, type = "histogram", size = 3, xparams = list(bins = 20, fill = "blue"), yparams = list(bins = 100, fill = "orange"))

#Create longer pivot for histogram plotting and rename
KSEDSlideGPDataLongLonger <- KSEDSlideGPDataLong %>%    
  
  pivot_longer(
    cols = -slide_name,                       # pivot all columns except case_id (all the symptoms columns)
    names_to = "annotation_type",             # assign name for new column that holds the symptoms
    values_to = "percent_gp") #%>%  # assign name for new column that holds the values (yes/no)
  
  #mutate(percent_gp = replace_na(percent_gp, "unknown")) # convert NA to "unknown"
head(KSEDSlideGPDataLongLonger)
KSEDSlideGPDataLongLonger$percent_gp<- as.numeric(as.character(KSEDSlideGPDataLongLonger$percent_gp))

# Step 1: Call the pdf command to start the plot
pdf(file = "./plots/HistoVisualEDvsWSI.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 6) # The height of the plot in inches
# Step 2: Create the plot with R code
ggplot(data = KSEDSlideGPDataLongLonger, mapping = aes(x = percent_gp, fill = annotation_type)) +
  geom_histogram(bins = 200, position = "dodge") +
  scale_y_sqrt() + #sqrt adjustment of y Axis
  labs(title = "Frequency distribution") + 
  geom_hline(yintercept=0, linetype = 2, color = "darkgray")
# Step 3: Run dev.off() to create the file!
dev.off()

ggplot(data = KSEDSlideGPDataLongLonger, mapping = aes(x = percent_gp, fill = annotation_type)) +
  geom_histogram(bins = 200, position = "dodge") +
  scale_y_sqrt() + #sqrt adjustment of y Axis
  labs(title = "Frequency distribution") + 
  geom_hline(yintercept=0, linetype = 2, color = "darkgray")

```


```{r, include=FALSE}
# Intraclass correlation coefficient (ICC) by GP

# Intraclass correlation coefficient (ICC) - Overall
icc(select(KSPeterSlideGPDataLong, percent_gp_KS, percent_gp_Peter), model = "oneway", type = "consistency", unit = "single")
# for Evelina/Danyil vs WSI
icc(select(KSEDSlideGPDataLong, percent_gp_KS, percent_gp_EvelinaDanyil), model = "oneway", type = "consistency", unit = "single")

# Intraclass correlation coefficient (ICC) - Desmo
KSSlideDesmo <- KSSlideDataWide %>% select(slide_name, desmo) %>% rename(desmo_KS = "desmo")
PeterSlideDesmo <- PeterVEstSlideDataWide %>% select(slide_name, desmo) %>% rename(desmo_Peter = "desmo")
KSPeterDesmo <- merge(KSSlideDesmo, PeterSlideDesmo)
head(KSPeterDesmo)

icc(select(KSPeterDesmo, desmo_KS, desmo_Peter), model = "oneway", type = "consistency", unit = "single")

# Intraclass correlation coefficient (ICC) - Replacement
KSSlideReplacement <- KSSlideDataWide %>% select(slide_name, replacement) %>% rename(replacement_KS = "replacement")
PeterSlideReplacement <- PeterVEstSlideDataWide %>% select(slide_name, replacement) %>% rename(replacement_Peter = "replacement")
KSPeterReplacement <- merge(KSSlideReplacement, PeterSlideReplacement)
head(KSPeterReplacement)

icc(select(KSPeterReplacement, replacement_KS, replacement_Peter), model = "oneway", type = "consistency", unit = "single")

# Intraclass correlation coefficient (ICC) - Pushing
KSSlidePushing <- KSSlideDataWide %>% select(slide_name, pushing) %>% rename(pushing_KS = "pushing")
PeterSlidePushing <- PeterVEstSlideDataWide %>% select(slide_name, pushing) %>% rename(pushing_Peter = "pushing")
KSPeterPushing <- merge(KSSlidePushing, PeterSlidePushing)
head(KSPeterPushing)

icc(select(KSPeterPushing, pushing_KS, pushing_Peter), model = "oneway", type = "consistency", unit = "single")
```

Probe categorization KS
```{r}
# Average GP %s for probes
KSProbeDataLong <- KSSlideDataLong %>% group_by(ids, annotation_types) %>% summarise(percent_gp = round(mean(percent_gp),2))
head(KSProbeDataLong)

# Categorize by GP predo
KSProbePredo <- KSProbeDataLong %>% group_by(ids) %>% slice_max(percent_gp) %>% select(ids, annotation_types) %>% rename(gp_predo = "annotation_types")
head(KSProbePredo)

# Filter out desmo values
KSProbeDesmo <- KSProbeDataLong %>% filter(annotation_types == "desmo")

# Categorize by 100 % desmo
KSProbeDesmo <- KSProbeDesmo %>% ungroup() %>% mutate(desmo_100 = cut(KSProbeDesmo$percent_gp, breaks = c(0, 99.9, 100), include.lowest = TRUE))

# Categorize by 95 % desmo
KSProbeDesmo <- KSProbeDesmo %>% ungroup() %>% mutate(desmo_95 = cut(KSProbeDesmo$percent_gp, breaks = c(0, 94.9, 100), include.lowest = TRUE))
head(KSProbeDesmo)

# Merge into dataset of probe ids and GP category groups
KSProbeCat <- merge(KSProbePredo, select(KSProbeDesmo, ids, desmo_100, desmo_95))
head(KSProbeCat)
```


```{r, include=FALSE}
# Probe categorization Peter

# Obs, we don't have GP lengths, only %s. Only ay to summarize by probe is averaging among slides. And same has to be done in KS data
PeterProbeData <- PeterVEstSlideGPDataLong %>% separate(slide_name, into = c("ids", "tumors", "blocks", "annotation_types"), sep = "-") %>% select(ids, annotation_types, percent_gp_Peter) %>% group_by(ids, annotation_types) %>% summarise(percent_gp_Peter = round(mean(percent_gp_Peter),2))
head(PeterProbeData)

# Categorize by GP predo
PeterProbePredo <- PeterProbeData %>% group_by(ids) %>% slice_max(percent_gp_Peter) %>% select(ids, annotation_types) %>% rename(gp_predo = "annotation_types")
head(PeterProbePredo)

#check that each ID has 1 value left
table(PeterProbePredo$ids)
table(table(PeterProbePredo$ids))

# DEBUG: probes with null % gps Peter
unique(PeterProbeData %>% filter(is.na(percent_gp_Peter)) %>% pull(ids))

# Filter out desmo values
PeterProbeDesmo <- PeterProbeData %>% filter(annotation_types == "desmo")

# Categorize by 100 % desmo
PeterProbeDesmo <- PeterProbeDesmo %>% ungroup() %>% mutate(desmo_100 = cut(PeterProbeDesmo$percent_gp_Peter, breaks = c(0, 99.9, 100), include.lowest = TRUE))

# Categorize by 95 % desmo
PeterProbeDesmo <- PeterProbeDesmo %>% ungroup() %>% mutate(desmo_95 = cut(PeterProbeDesmo$percent_gp_Peter, breaks = c(0, 94.9, 100), include.lowest = TRUE))
head(PeterProbeDesmo)

# Merge into dataset of probe ids and GP category groups
PeterProbeCat <- merge(PeterProbePredo, select(PeterProbeDesmo, ids, desmo_100, desmo_95))
head(PeterProbeCat)
```

Probe categorization Evelina/Danyil
```{r}
# Summarise calculates the mean of each slide per ID, so that we have the predominant growth pattern per patient afterwards
EDProbeData <- EDEstSlideDataLong %>% separate(slide_name, into = c("ids", "tumors", "blocks", "annotation_types"), sep = "-") %>% select(ids, annotation_types, percent_gp_EvelinaDanyil) %>% group_by(ids, annotation_types) %>% summarise(percent_gp_EvelinaDanyil = round(mean(percent_gp_EvelinaDanyil),2))
head(EDProbeData)

#check that each ID has 3 values
table(EDProbeData$ids)
table(table(EDProbeData$ids))

# Categorize by GP predo
EDProbePredo <- EDProbeData %>% group_by(ids) %>% slice_max(percent_gp_EvelinaDanyil) %>% select(ids, annotation_types) %>% rename(gp_predo = "annotation_types")
head(EDProbePredo)

#check that each ID has 1 value left
table(EDProbePredo$ids)
table(table(EDProbePredo$ids))

# DEBUG: How to handle slides with 50% growth pattern


# DEBUG: probes with null % gps Evelina/Danyil
unique(EDProbeData %>% filter(is.na(percent_gp_EvelinaDanyil)) %>% pull(ids))

# Filter out desmo values
EDProbeDesmo <- EDProbeData %>% filter(annotation_types == "desmo")

# Categorize by 100 % desmo
EDProbeDesmo <- EDProbeDesmo %>% ungroup() %>% mutate(desmo_100 = cut(EDProbeDesmo$percent_gp_EvelinaDanyil, breaks = c(0, 99.9, 100), include.lowest = TRUE))

# Categorize by 95 % desmo
EDProbeDesmo <- EDProbeDesmo %>% ungroup() %>% mutate(desmo_95 = cut(EDProbeDesmo$percent_gp_EvelinaDanyil, breaks = c(0, 94.9, 100), include.lowest = TRUE))
head(EDProbeDesmo)

# Merge into dataset of probe ids and GP category groups
EDProbeCat <- merge(EDProbePredo, select(EDProbeDesmo, ids, desmo_100, desmo_95))
head(EDProbeCat)
```


```{r, include=FALSE}
# Cohen's Kappa Peter

# GP predo
KSPeterPredo <- merge(KSProbeCat %>% select(ids, gp_predo) %>% rename(gp_predo_KS = "gp_predo"), PeterProbeCat %>% select(ids, gp_predo) %>% rename(gp_predo_Peter = "gp_predo"))
kappa2(select(KSPeterPredo, gp_predo_KS, gp_predo_Peter))

# desmo 100
KSPeterDesmo100 <- merge(KSProbeCat %>% select(ids, desmo_100) %>% rename(desmo_100_KS = "desmo_100"), PeterProbeCat %>% select(ids, desmo_100) %>% rename(desmo_100_Peter = "desmo_100"))
kappa2(select(KSPeterDesmo100, desmo_100_KS, desmo_100_Peter))

# desmo 95
KSPeterDesmo95 <- merge(KSProbeCat %>% select(ids, desmo_95) %>% rename(desmo_95_KS = "desmo_95"), PeterProbeCat %>% select(ids, desmo_95) %>% rename(desmo_95_Peter = "desmo_95"))
kappa2(select(KSPeterDesmo95, desmo_95_KS, desmo_95_Peter))
```

Cohen's Kappa Evelina/Danyil visual scoring versus annotation-based scoring
```{r}
# GP predo
KSEDPredo <- merge(KSProbeCat %>% select(ids, gp_predo) %>% rename(gp_predo_KS = "gp_predo"), EDProbeCat %>% select(ids, gp_predo) %>% rename(gp_predo_EvelinaDanyil = "gp_predo"))
head(KSEDPredo)
table(table(KSEDPredo$ids))
kappa2(select(KSEDPredo, gp_predo_KS, gp_predo_EvelinaDanyil))

# desmo 100
KSEDDesmo100 <- merge(KSProbeCat %>% select(ids, desmo_100) %>% rename(desmo_100_KS = "desmo_100"), EDProbeCat %>% select(ids, desmo_100) %>% rename(desmo_100_ED = "desmo_100"))
head(KSEDDesmo100)
table(table(KSEDDesmo100$ids))
kappa2(select(KSEDDesmo100, desmo_100_KS, desmo_100_ED))

# desmo 95
KSEDDesmo95 <- merge(KSProbeCat %>% select(ids, desmo_95) %>% rename(desmo_95_KS = "desmo_95"), EDProbeCat %>% select(ids, desmo_95) %>% rename(desmo_95_ED = "desmo_95"))
head(KSEDDesmo95)
table(table(KSEDDesmo95$ids))
kappa2(select(KSEDDesmo95, desmo_95_KS, desmo_95_ED))
```